### 2015-06-28 narrative/intro.tex
working more on the introduction

\section*{Introduction} {\color{red} Scott, Hasan, Gary, Jay - 4 pages}
\label{sec:introduction}
The Lightweight File Systems~\cite{lwfs} (LWFS) project from Sandia sought to
demonstrate stripping down a file system to the bare minimum features required
for any storage system. Then, additional functionality and semantics, such as
POSIX-style directories and consistency requirements, are added in using
auxiliary services separate from the stack or by layering providing a new
interface to the storage system. The project demonstrated that using a core
object storage system with integrated security providing authentication and
autorization services, customized storage interfaces could be developed with
exactly the overhead required for different applications. The second phase of
LWFS, known as Sirocco~\cite{sirocco}, is investigating how to manage storage
devices for efficient and scalable storage performance.  Sirocco has been in
process for three years and has a usable storage infrastructure system. This
proposal seeks to build upon this work to demonstrate an application, data, and
usage aware storage system optimized for exascale composite applications and
capability application runs. We are proposing to combine the applications
knowledge from ORNL with the SNL LWFS system and storage system knowledge, the
metadata and storage systems knowledge of UCSC, and the middleware expertise of
Rutgers to investigate how to make a data, workflow, and application aware
storage system.

From the perspective of the simulation scientist, they want to be able to
have knowledge about how much time they will put into writing their data,
and how much storage space they will be able to get during the lifetime of
their simulations. Furthermore, they would like the ability to get a certain
amount of Quality of Service such that they can then make decisions when the
expected bandwidth, for example is less than what they desire. The user will
then place in a certain set of rules which the system can make autonomic
decisions to help decide what should be done. In our simple scenario, a user
wants to perhaps write 1 PB of data every hour for their checkpoint/restart,
and they which to write 500TB of data every 15 minutes for their analysis
and visualization results which have been reduced by using in situ reduction
techniques. At this stage, they will ask the system to write their data, and
they would like to get an estimated time which they can then figure out,
through a series of rules, whether they want to write out all of the data or
if they want to write none of the data but rather wait for a later time to
write or write a reduced amount of data, along with some code container to
allow them to regenerate the data with a certain level of accuracy.
Furthermore, they realize that the analysis data will be saved on either the
parallel file system, or the campaign storage, a longer term storage area
with lower bandwidth than the parallel file system, or if some of the data
will just go to tape. Each piece of data the user will specify the time they
will want to keep the data round. We want the system to give the users a
certain amount of currency in terms of bandwidth, storage space on each
level, and latency expectations. These notions will be fuzzy but they will
allow the user to make ad-hoc decisions to figure out what needs to be
saved.

From the perspective of the storage system, it needs to manage the data from
a given user request amongst all of the request and try to optimize the
entire system. Today the problem is that when a user writes there is a
tremendous amount of interference from other users on the system who may be
writing or reading at the same time. By maiming a certain level of quality
of service the system must potentially lock out users with lower currency
they want to place in the reading or writing of their data. The storage
system must also automatically move data from one tier of the storage to
another, without effecting the quality of service that it guarantees at the
time of questioning. The storage system must also look at ways to organize
data amongst all of the tiers of storage in one consistent way.

From the perspective of the users who want to read in the data from their
simulations, and then operate on this data and possibly write data, we see
that in general these users will be running on much less compute resources
than the users who are running the simulation. One of the important aspects
of this class of user is that they want a reduction of latency and they
require a certain amount of quality of service as well. When someone is
doing interactive analysis and expects that the data they read in is about
10s, but it terms into 1,000s, the user typically tends to either wait
another day for doing their analysis, or gets frustrated and reads much less
data than they really need to. Furthermore the user tends to read in just a
small sample of the data without much knowledge if they are missing
important pieces of the data.

Our research into the storage system and middle ware layers must address
many of these challenges including
\begin{enumerate}
\item How do we describe the user intentions at the API layer and have this
  communicated down to the middle-ware and storage layers?
\item How do we allow users to define their user defined compression
  techniques and have this data read back from the storage system layers?
  Where do we execute the decompression techniques 
\item How do we evaluate the tradeoffs at runtime to guide data placement,
  including the movement of data across the network, across the different
  storage tiers? (data movement and quality tradeoff)
\item Can we use different forms of learning techniques as daemons on the
  system to perform data migration from one storage tier to another? For
  example, if we see that a user is looking at one time slice after another
  for a certain object in their dataset which is in the slower tiers of the
  hierarchy, do we automatically propane up the data which has not been
  requested in the hope that this will be requested data? (prediction and
  prefetching, user defined compression on movement)
\item Can we understand the true need of campaign storage and understand the
  possible impact of the campaign storage if we run a hadoop-like
  file-system instead of a lustre/gpfs file system? (additional layers)
\item Can we limit the amount of data duplication? Data space is going to be
  very constrained on the exascale system so we must ensure that minimal
  copies are made of data, and when data is duplicated, they are removed
  through a garbage collection routine. (space management)
\item Can we build a model to give us the time estimates during the reading
  and writing phases, so that rules can be in place from the user
  perspective to make adaptive decisions. For example, if the file system
  says that reading will take 3 months, users can they place in rules to
  then read in a subset of data, and they can understand which data can be
  read quickly and which pieces will take more time. (estimation)
\item Can we understand how to place code in the system which will allow
  data-regeneration to take place. (regeneration)
\item How does the storage system do a better job in managing request from
  all of the users on a LCF than today? We realize that today users who have
  a better middlware system can often lock other users from getting high
  performance when they are running. If we have the concept of currency
  which is eventually used in the same extent as node-hours, then users will
  have to be able to think about how much storage and how much bandwidth
  they can choose. One question that needs to be understood is if there are
  times when the system sees that there are very few storage system
  resources being used, then the lucky users can get the bandwidth cheaper
  than at times of heavy usage. How can we enforce this? (Fairness)
\item How do we manage all of the metdata not only from the principle
  objects, but from the sub-objects? How well will this scale when we have
  users who can potentially create billions of objects from their
  simulation? (scalability and discovery)
\end{enumerate}

@@

We will achieve reduced time to knowledge using a combination of techniques;
\begin{enumerate}
\item Data annotations specified at the application level to quantify the
  relative important, utility and lifetime of data objects;
\item Partitioning of data objects across the storage hierarchy utilizing
  the additional knowledge embedded in the annotations;
\item Evaluating the tradeoffs at runtime to guide data placement, movement,
  and migration across storage layers using models, heuristics and
  continuous learning;
\item Utilize the additional knowledge available about the data to perform
  application-aware data compression and I/O prioritization;
\end{enumerate}

We are aiming to spread an output across the vertical layers of the storage
hierarchy simultaneously. When data should migrate to a higher tier, what
happens to the existing version?

Is there a canonical copy that is the originally written version?

Is there annotation about this (I think so) attached to that part of the
variable?

At the tape layer, we see having a lossy compressed version just above the
tape layer used as a directory/index and then the ``full'' resolution
version on tape only retrieving once the user accepts the time/quality
tradeoff. This is going to require some serious language to describe. If we
have 100s PB of tape space, we'll need 1s PB of disk even at a 99\% data
reduction. That is non-trivial.

What happens when that data is retrieved from tape? Is a third copy made in
an appropriately sized/quality level version for the tier it is requested to
be pulled to? What if it is just the index version? Or the on tape version?
Does it have a TTL because of the pain retrieving it from tape?

Do we specify a target tier or just allow the system to place it in a place
that makes sense? Given the potential space limitations, I think this is
pretty critical because it could cause evictions or other insufficient space
actions that are unintended. I can see us specifying that when you ask for
data to be pulled at a particular quality level to a particular performance
tier that it is a best effort with the use specifying if compromise or just
failure is the result if the request cannot be fulfilled. Making this
interaction make sense is going to require some reasonable thought and
testing. We'll probably need to run it past apps people to get their
feedback too as I don't think we are able to give a solid answer without
some broad input.

How does the migration to ``data lakes''/``campaign storage'' work? How does
the migration to tape ultimately happen? I think we need to incorporate some
explicit staging commands to move data both up and down the stack along with
some variability in placement based on data features and current system
state.

Given the storage scarcity, particularly for NVM, I think we need to have a
solid story here as part of the proposal. Yes, there are questions that have
to be answered to build this still, but we need to have some pretty solid
clarity where possible. I don't feel like I can explain it well enough at
what I believe to be a proper clarity level at this point.

>>>>

\section*{Introduction} {\color{red} Scott, Hasan, Gary, Jay - 4 pages}
\label{sec:introduction}

@@

To address these challenges we will We are proposing to combine the applications
knowledge from ORNL, from the ADIOS project, with the SNL LWFS system and storage system knowledge, the
metadata and storage systems knowledge of UCSC, and the middleware expertise of
Rutgers to investigate how to make a data, workflow, and application aware
storage system.

%ADIOS was developed with the understanding that we must
%not only address the bottlenecks for current applications and
%hardware platforms but also provide a path forward for the next
%generation of applications and systems that would need to both
%maximize bandwidth to the storage system and also support
%transparently working around the storage system bandwidth
%limitations with new techniques and tools. To support the
%diverse operating modes of both using persistent storage and
%other data storage and processing technology, we made a
%great effort to provide a simplified interface to application
%developers, offering a simple, portable and scalable way for
%scientists to manage data that may need to be written, read
%or processed during simulation runs. This required abstracting
%away many decisions typically made in the application code
%so that they may be configured externally.
%In addition to this abstracted interface with external configuration
%options, common services were incorporated to afford
%optimizations beyond those for a single platform, such as
%buffering, aggregation, subfiling, and chunking with options
%to select each based on the data distribution characteristics
%of the application. A variety of asynchronous I/O techniques
%have been investigated and are being integrated with ADIOS.
%A recognition of application complexity has led to new techniques
%for testing I/O performance by extracting I/O patterns
%from applications and automatically generating benchmark
%codes. And finally, the march toward exascale has fueled the
%need to consider additional features such as compression and
%indexing to better cope with the expected tsunami of data.

In this project we extend ADIOS with new APIs and novel techniques to talk to the
storage back-end. We will also test our ideas on the next generation OLCF system along with
the next generation NERSC system. One of the clear challenges of this work will be in accessing new APIs which
will give the applications much more control and flexibility of their I/O.
%
In this project we will make use of the many inovations we discovered in 
the Lightweight File Systems~\cite{lwfs} (LWFS) project, where Sandia sought to
demonstrate stripping down a file system to the bare minimum features required
for any storage system. Then, additional functionality and semantics, such as
POSIX-style directories and consistency requirements, are added in using
auxiliary services separate from the stack or by layering providing a new
interface to the storage system. The project demonstrated that using a core
object storage system with integrated security providing authentication and
authorization services, customized storage interfaces could be developed with
exactly the overhead required for different applications. The second phase of
LWFS, known as Sirocco~\cite{sirocco}, is investigating how to manage storage
devices for efficient and scalable storage performance.  Sirocco has been in
process for three years and has a usable storage infrastructure system.

%  This
%proposal seeks to build upon this work to demonstrate an application, data, and
%usage aware storage system optimized for exascale composite applications and
%capability application runs. We are proposing to combine the applications
%knowledge from ORNL with the SNL LWFS system and storage system knowledge, the
%metadata and storage systems knowledge of UCSC, and the middleware expertise of
%Rutgers to investigate how to make a data, workflow, and application aware
%storage system.

\subsection{Writing Perspective}
\label{subsec:sim-perspective}
From the perspective of the simulation scientist, they want to be able to
have knowledge about how much time they will put into writing their data,
and how much storage space they will be able to get during the lifetime of
their simulations. Furthermore, they would like the ability to get a certain
amount of Quality of Service such that they can then make decisions when the
expected bandwidth, for example is less than what they desire. The user will
then place in a certain set of rules which the system can make autonomic
decisions to help decide what should be done.
%

One of the key aspects of our approach is to allow the users to define the utility (importance) of their data;
in other-words, for each data object, users need to come up with a way to define pieces of the object
which have more utility (more importance) . There are many clear research challenges faced with this approach 
including: 1) How do we make this simple enough so that users can do this, 2) can we come up with a model to
allow the system to then make the autonomic decisions to save the data which has the highest utility, and 3) can
we allow the back-end storage system to interact with this definition so that it can optimize not just over one application
but across all of the applications.
 %
  In our simple scenario, a user
wants to perhaps write 1 PB of data every hour for their checkpoint/restart,
and they which to write 500TB of data every 15 minutes for their analysis
and visualization results which have been reduced by using in situ reduction
techniques. At this stage, they will ask the system to write their data, and
they would like to get an estimated time which they can then figure out,
through a series of rules, whether they want to write out all of the data or
if they want to write none of the data but rather wait for a later time to
write or write a reduced amount of data, along with some code container to
allow them to regenerate the data with a certain level of accuracy.
Furthermore, they realize that the analysis data will be saved on either the
parallel file system, or the campaign storage, a longer term storage area
with lower bandwidth than the parallel file system, or if some of the data
will just go to tape. Each piece of data the user will specify the lifetime so that the 
data with the most important utility is kept around for the longest period of time, but the data which has the least utility might
be migrated to lower layers of the storage hierarchy faster.

\subsection{Storage System Perspective}
\label{subsec:storage-perspective}
From the perspective of the storage system, it needs to manage the data from
a given user request amongst all of the request and try to optimize the
entire system. Today the problem is that when a user writes there is a
tremendous amount of interference from other users on the system who may be
writing or reading at the same time. By maiming a certain level of quality
of service the system must potentially lock out users with lower currency
they want to place in the reading or writing of their data. The storage
system must also automatically move data from one tier of the storage to
another, without effecting the quality of service that it guarantees at the
time of questioning. The storage system must also look at ways to organize
data amongst all of the tiers of storage in one consistent way.

One of the clear research objectives of this project is to understand how the storage system can interact
with the middleware (application-aware) piece so that the data with the highest utility is kept on the fastest 
layers of the storage system for as long as specified. Clearly we need to maintain some sort of QoS such that
we can give fuzzy-estimates of the time it will take to write/read data from the different layers, and clearly it is critical
that the storage system can manage the data across all of the layers. We will use learning techniques to automatically
migrate data sub-objects across the hierarchy as well as to eventually evict data which is large and has very low utility.

\subsection{Reading  Perspective}
\label{subsec:reading-perspective}
From the perspective of the users who want to read in the data from their
simulations, and then operate on this data and possibly write data, we see
that in general these users will be running on much less compute resources
than the users who are running the simulation. One of the important aspects
of this class of user is that they want a reduction of latency and they
require a certain amount of quality of service as well. When someone is
doing interactive analysis and expects that the data they read in is about
10s, but it terms into 1,000s, the user typically tends to either wait
another day for doing their analysis, or gets frustrated and reads much less
data than they really need to. Furthermore the user tends to read in just a
small sample of the data without much knowledge if they are missing
important pieces of the data.

Our research into the storage system and middle ware layers must address
many of these challenges including
\begin{enumerate}
\item How do we describe the user intentions at the API layer and have this
  communicated down to the middle-ware and storage layers?
\item How do we allow users to define their user defined compression
  techniques and have this data read back from the storage system layers?
  Where do we execute the decompression techniques 
\item How do we evaluate the tradeoffs at runtime to guide data placement,
  including the movement of data across the network, across the different
  storage tiers? (data movement and quality tradeoff)
\item Can we use different forms of learning techniques as daemons on the
  system to perform data migration from one storage tier to another? For
  example, if we see that a user is looking at one time slice after another
  for a certain object in their dataset which is in the slower tiers of the
  hierarchy, do we automatically propane up the data which has not been
  requested in the hope that this will be requested data? (prediction and
  prefetching, user defined compression on movement)
\item Can we understand the true need of campaign storage and understand the
  possible impact of the campaign storage if we run a hadoop-like
  file-system instead of a lustre/gpfs file system? (additional layers)
\item Can we limit the amount of data duplication? Data space is going to be
  very constrained on the exascale system so we must ensure that minimal
  copies are made of data, and when data is duplicated, they are removed
  through a garbage collection routine. (space management)
\item Can we build a model to give us the time estimates during the reading
  and writing phases, so that rules can be in place from the user
  perspective to make adaptive decisions. For example, if the file system
  says that reading will take 3 months, users can they place in rules to
  then read in a subset of data, and they can understand which data can be
  read quickly and which pieces will take more time. (estimation)
\item Can we understand how to place code in the system which will allow
  data-regeneration to take place. (regeneration)
\item How does the storage system do a better job in managing request from
  all of the users on a LCF than today? We realize that today users who have
  a better middlware system can often lock other users from getting high
  performance when they are running. If we have the concept of currency
  which is eventually used in the same extent as node-hours, then users will
  have to be able to think about how much storage and how much bandwidth
  they can choose. One question that needs to be understood is if there are
  times when the system sees that there are very few storage system
  resources being used, then the lucky users can get the bandwidth cheaper
  than at times of heavy usage. How can we enforce this? (Fairness)
\item How do we manage all of the metdata not only from the principle
  objects, but from the sub-objects? How well will this scale when we have
  users who can potentially create billions of objects from their
  simulation? (scalability and discovery)
\end{enumerate}
%
%Exascale scientific discovery will be severely bottlenecked without
%sufficient new research into managing and storing the large amounts of data
%that will be produced during the simulation, and analyzed for months
%afterwards.
%%
%In this project we will demonstrate novel techniques to facilitate efficient
%mapping of data objects, even partitioning individual variables, from the
%user space onto multiple storage tiers, and enable application-guided data
%reductions/transformations to address capacity and bandwidth bottlenecks,
%while constraining the error to be within user provided bounds.
%%
%We will address the associated I/O and storage challenges in the context of
%current and emerging storage landscapes, and expedite insights into critical
%scientific processes, demonstrating the validity of our approach in key DOE
%domains. Our techniques will be to research novel techniques into a Scalable
%Storage Software Infrastructure by integrating services from the middle-ware
%layer which will talk to the applications, with the Storage system. The
%negotation between these layers will be a fundamental service which we will
%create in this project to ensure that users will be able to save `'the
%best'' amount of data in the different storage tiers.
%
%The metric we are most interested in optimizing is time to knowledge.
%Current approaches to addressing the I/O bottleneck fall into two broad
%categories. Parallel file system approaches that optimize the throughput for
%an entire system, and I/O middleware approaches that optimize the
%performance of a single application. Both approaches have seen success but
%are unlikely to overcome the major obstacles in reaching exascale. Instead
%of trying to optimize throughput, we will seek to reduce the time to
%knowledge. This is the most significant metric for scientific applications
%where the desired outcome is not storing data, but rather executing a
%knowledge extraction process on the data. Moreover, we aim to perform the
%optimization not for a single application, but rather for workload of a
%multi-user multi-application system environment.
%
%Our approach leverages the expected characteristics of exascale storage
%hardware. The storage layer will be partitioned into multiple heterogenous
%tiers with vastly different performance characteristics. This difference
%between layers will be further exacerbated by the constraints on capacity
%and data lifetime within a layer. Tape archival storage will still maintain
%data long term, but access to this data will be orders of magnitude slower
%than the next layer and naively accessing data from archival storage will
%greatly impact productivity.
%
%We will achieve reduced time to knowledge using a combination of techniques;
%\begin{enumerate}
%\item Data annotations specified at the application level to quantify the
%  relative important, utility and lifetime of data objects;
%\item Partitioning of data objects across the storage hierarchy utilizing
%  the additional knowledge embedded in the annotations;
%\item Evaluating the tradeoffs at runtime to guide data placement, movement,
%  and migration across storage layers using models, heuristics and
%  continuous learning;
%\item Utilize the additional knowledge available about the data to perform
%  application-aware data compression and I/O prioritization;
%\end{enumerate}
%
%We are aiming to spread an output across the vertical layers of the storage
%hierarchy simultaneously. When data should migrate to a higher tier, what
%happens to the existing version?
%
%Is there a canonical copy that is the originally written version?
%
%Is there annotation about this (I think so) attached to that part of the
%variable?
%
%At the tape layer, we see having a lossy compressed version just above the
%tape layer used as a directory/index and then the ``full'' resolution
%version on tape only retrieving once the user accepts the time/quality
%tradeoff. This is going to require some serious language to describe. If we
%have 100s PB of tape space, we'll need 1s PB of disk even at a 99\% data
%reduction. That is non-trivial.
%
%What happens when that data is retrieved from tape? Is a third copy made in
%an appropriately sized/quality level version for the tier it is requested to
%be pulled to? What if it is just the index version? Or the on tape version?
%Does it have a TTL because of the pain retrieving it from tape?
%
%Do we specify a target tier or just allow the system to place it in a place
%that makes sense? Given the potential space limitations, I think this is
%pretty critical because it could cause evictions or other insufficient space
%actions that are unintended. I can see us specifying that when you ask for
%data to be pulled at a particular quality level to a particular performance
%tier that it is a best effort with the use specifying if compromise or just
%failure is the result if the request cannot be fulfilled. Making this
%interaction make sense is going to require some reasonable thought and
%testing. We'll probably need to run it past apps people to get their
%feedback too as I don't think we are able to give a solid answer without
%some broad input.
%
%How does the migration to ``data lakes''/``campaign storage'' work? How does
%the migration to tape ultimately happen? I think we need to incorporate some
%explicit staging commands to move data both up and down the stack along with
%some variability in placement based on data features and current system
%state.
%
%Given the storage scarcity, particularly for NVM, I think we need to have a
%solid story here as part of the proposal. Yes, there are questions that have
%to be answered to build this still, but we need to have some pretty solid
%clarity where possible. I don't feel like I can explain it well enough at
%what I believe to be a proper clarity level at this point.
